---
title: "Peer Assessment I"
output:
  html_document: 
    pandoc_args: [
      "--number-sections",
    ]
---


First, let us load the data and necessary packages:

```{r load, message = FALSE}
load("ames_train.Rdata")
library(MASS)
library(dplyr)
library(ggplot2)
```

#
Make a labeled histogram (with 30 bins) of the ages of the houses in the data set, and describe the distribution.


```{r Q1}
ames_train %>% summarise(n_distinct(Year.Built)) #102 distinct values
# Create a histogram with 30 bins of Year.Built
ggplot(data = ames_train, aes(Year.Built)) + geom_histogram(binwidth = (102/30))

```

* * *
The data appear to be left skewed, and multi-modal.

* * *


#
The mantra in real estate is "Location, Location, Location!" Make a graphical display that relates a home price to its neighborhood in Ames, Iowa. Which summary statistics are most appropriate to use for determining the most expensive, least expensive, and most heterogeneous (having the most variation in housing price) neighborhoods? Report which neighborhoods these are based on the summary statistics of your choice. Report the value of your chosen summary statistics for these neighborhoods.


```{r Q2}
ggplot(data = ames_train, aes(Neighborhood, price)) + geom_boxplot() + coord_flip()
summary_stats <- 
  ames_train %>% 
  filter(!is.na(price)) %>%
  group_by(Neighborhood) %>% 
  summarise(min = min(price), max = max(price), mean = mean(price), median = median(price), sd = sd(price))
```


* * *

The Oldtown Neighborhood had the lowest priced home, at \$12,789, and the fourth lowest mean (\$120,225) and median (\$120,000) values for all houses in the area.
The highest priced home was in Northridge Heights, at \$615,000. This Northridge heights had the second highest mean and median values, as well.
The neighborhood with the most price diversity was Stonebridge, with a standard deviation of \$123,459. Stonebridge has the highest mean and median home prices, as well.


* * *

# 

Which variable has the largest number of missing values? Explain why it makes sense that there are so many missing values for this variable.

```{r Q3}
sapply(ames_train, function(x) sum(is.na(x)))

```


* * *

Pool.QC seems to have the most missing data (997 cases), which makes sense, as this value would only be filled if the property has a pool.


* * *

#

We want to predict the natural log of the home prices. Candidate explanatory variables are lot size in square feet (Lot.Area), slope of property (Land.Slope), original construction date (Year.Built), remodel date (Year.Remod.Add), and the number of bedrooms above grade (Bedroom.AbvGr). Pick a model selection or model averaging method covered in the Specialization, and describe how this method works. Then, use this method to find the best multiple regression model for predicting the natural log of the home prices.


```{r Q4}
# Need to natural log Lot.Area to make the relationship more linear
ggplot(data = ames_train, aes(Lot.Area, log(price))) + geom_point() 

# Need to deal with skew in year.built...
# Need to deal with
ggplot(data = ames_train, aes(Bedroom.AbvGr, log(price))) + geom_point() 

       
model_base <- lm(data = ames_train, log(price) ~ Lot.Area + Land.Slope + Year.Built + Year.Remod.Add + Bedroom.AbvGr)


model2 <- model2 <- lm(data = ames_train, log(price) ~ Lot.Area + Year.Built + Year.Remod.Add + Bedroom.AbvGr)
model3 <- model2 <- lm(data = ames_train, log(price) ~ Lot.Area + Land.Slope + Year.Remod.Add + Bedroom.AbvGr)
model4 <- model2 <- lm(data = ames_train, log(price) ~ Lot.Area + Land.Slope + Year.Built + Bedroom.AbvGr)
model5 <- model2 <- lm(data = ames_train, log(price) ~ Lot.Area + Land.Slope + Year.Built + Year.Remod.Add)
model6 <- lm(data = ames_train, log(price) ~ Land.Slope + Year.Built + Year.Remod.Add + Bedroom.AbvGr)
model_results <- data.frame(c("model_base", "model2", "model3", "model4", "model5", "model6"), c(summary(model_base)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared, summary(model4)$adj.r.squared, summary(model5)$adj.r.squared, summary(model6)$adj.r.squared))
names(model_results) <- c("model", "adj.r.squared")
model_results
summary(model_base)
```

* * *

We'll use a backwards step-wise selection approach to selecting a multiple linear regression. It finds that all variables are significant, and contribute to a higher adjusted R^2 value.


* * *

#

Which home has the largest squared residual in the previous analysis (Question 4)? Looking at all the variables in the data set, can you explain why this home stands out from the rest (what factors contribute to the high squared residual and why are those factors relevant)?


```{r Q5}
plot(model_base$residuals)

```

* * *

NOTE: Write your written response to Question 5 here.  Delete this note before you submit your work.


* * *

#

Use the same model selection method you chose in Question 4 to again find the best multiple regression model to predict the natural log of home prices, but this time **replacing Lot.Area with log(Lot.Area)**. Do you arrive at a model including the same set of predictors?


```{r Q6}
model_base <- lm(data = ames_train, log(price) ~ log(Lot.Area) + Land.Slope + Year.Built + Year.Remod.Add + Bedroom.AbvGr)


model2 <- model2 <- lm(data = ames_train, log(price) ~ log(Lot.Area) + Year.Built + Year.Remod.Add + Bedroom.AbvGr)
model3 <- model2 <- lm(data = ames_train, log(price) ~ log(Lot.Area) + Land.Slope + Year.Remod.Add + Bedroom.AbvGr)
model4 <- model2 <- lm(data = ames_train, log(price) ~ log(Lot.Area) + Land.Slope + Year.Built + Bedroom.AbvGr)
model5 <- model2 <- lm(data = ames_train, log(price) ~ log(Lot.Area) + Land.Slope + Year.Built + Year.Remod.Add)
model6 <- lm(data = ames_train, log(price) ~ Land.Slope + Year.Built + Year.Remod.Add + Bedroom.AbvGr)
model_results <- data.frame(c("model_base", "model2", "model3", "model4", "model5", "model6"), c(summary(model_base)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared, summary(model4)$adj.r.squared, summary(model5)$adj.r.squared, summary(model6)$adj.r.squared))
names(model_results) <- c("model", "adj.r.squared")
model_results
summary(model_base)

```

* * *

#

Do you think it is better to log transform Lot.Area, in terms of assumptions for linear regression? Make graphs of the predicted values of log home price versus the true values of log home price for the regression models selected for Lot.Area and log(Lot.Area). Referencing these two plots, provide a written support that includes a quantitative justification for your answer in the first part of question 7.

```{r Q7}
ggplot(data = ames_train, aes())

```

* * *

NOTE: Write your written response to Question 7 here.  Delete this note before you submit your work.


* * *
###